Scattering from compact objects: \\ Regge poles and the Complex Angular Momentum method

### 2019-12-13 rainbowCAM.tex
First draft of conclusions, although there may be more to add.

The Regge pole spectrum for a compact body is qualitatively similar to the Regge pole spectrum found in Mie scattering of electromagnetic waves by a transparent droplet of refractive index $\tilde{n}$. This has been studied since the 1960s; see for example Fig.~9.2 in Ref.~\cite{Nussenzveig:2006}. % In the latter case, the poles are given by the zeros of a combination of special functions.
Henceforth we shall adopt the terminology of Nussenzveig \cite{Nussenzveig:2006}, in which the three branches are labelled as:
\begin{enumerate}
 \item \emph{Broad resonances}: approximately uniformly-spaced poles above the real axis with approximately constant imaginary part; somewhat sensitive to internal structure ($r < R$);
 \item \emph{Surface waves}: highly-damped modes that are relatively insensitive to the internal structure and which depend chiefly on the surface geometry.
 \item \emph{Narrow resonances}: modes approaching the real axis corresponding to trapped modes which only appear in the UCO case ($R < 3M$).
\end{enumerate}
In Figs.~\ref{fig:rp1} and \ref{fig:rp2} the broad resonances, surface waves and narrow resonances are indicated by red squares, blue circles and purple triangles, respectively; and black diamonds indicate the black hole RPs. Whereas there exists an infinite number of poles of the surface-wave branch, in principle, there are no narrow-resonance poles at all in the $R/M = 6$ case, and just $\le 4$ narrow-resonance poles in the $R/M = 2.26$ case. The narrow-resonance branch is observed to end close to the start of the black-hole branch.

@@

Regge poles and quasinormal modes for relativistic stellar models (of which \textit{w}-modes are a sub-category for gravitational perturbations) both satisfy the same wave equation and the same boundary conditions but with different interpretations for the angular momentum index and the frequency. Both types of pole satisfy the regularity condition at the origin~\eqref{bc_1_in} and the condition of a purely outgoing wave in the far field
%
\begin{equation}\label{bc_pole_inf}
\phiout (r) \scriptstyle{\underset{r_\ast \to +\infty}{\sim}}
\displaystyle{  A^{(+)}_{\lambda-1/2} (\omega) e^{+i\omega r_\ast}}.
\end{equation}
%
Thus, the Regge poles are solutions of Eq.~\eqref{H_Radial_equation} for which the  Wronskian of $\phireg$ and $\phiout$ vanishes (\textit{i.e.}, $A^{(-)}_{\lambda_n(\omega)-1/2} (\omega)=0$), viz.,
\begin{equation}\label{Wronskian}
  W[\phireg, \phiout ]= 0 .
\end{equation}
It has been shown \cite{Zhang:2011pq} that the asymptotic expressions of $\phireg$ and $\phiout$ can be derived in asymptotic regions, by using the standard WKB approximation. In the high-frequency limit, the interior solution of Eq.~(\ref{H_Radial_equation}) is (see Ref.~\cite{Zhang:2011pq})
\begin{equation}\label{Approx_origin}
  \phireg \underset{\omega \to \infty}{=} \omega r_* j_{\lambda-1/2} (\omega r_*),  \quad\quad 0\leq r_*\leq R_* ,
\end{equation}
where $j_{\lambda-1/2}(\cdot)$ is the spherical Bessel function of the first kind. In addition
\begin{equation}
\label{Approx_infinity}
\phiout  \scriptstyle{\underset{\omega \to \infty}{=}}
\left\{
\begin{aligned}
&e^{i\omega(r_*-R_*)}+\mathcal{R} e^{-i\omega (r_*-R_*)}& \scriptstyle{1/\omega\leq r_{*} < R_*,}\\
&\left(1+\mathcal{R}\right) e^{i\omega(r_*-R_*)}        & \scriptstyle{R_*\leq r_{*}<\infty.}
\end{aligned}
\right.
\end{equation}
Here $R_*$ is the tortoise coordinate at the surface of the body,
\begin{equation}\label{R_star}
  R_*=\int_{0}^{R}\, dr\,\frac{1}{\sqrt{f(r)h(r)}}
\end{equation}
and $\mathcal{R}$ is a reflection coefficient with the definition given in Ref.~\cite{Berry_1982}.
Because the potential has a direct discontinuity at the surface of the compact body (see. Refs \cite{Zhang:2011pq,Berry_1982} for more details), we have for our model
%
\begin{equation}\label{R_reflection_1}
  \mathcal{R} = \alpha \omega^{-2}
\end{equation}
with
\begin{eqnarray}
% \nonumber % Remove numbering (before each equation)
  \alpha &=& \frac{1}{4} \Delta V(R) = \pm \frac{3M(R-2M)}{4 R^4} ,
\end{eqnarray}
where $\Delta V$ is the discontinuity in the effective potential at the surface, defined in Eq.~(\ref{def:DeltaV}), and the $+$ ($-$) sign denotes the scalar (axial gravitational wave) case.

@@

\sam{Sam to write this}

In this work, we have calculated the full spectrum of Regge poles for a compact body spacetime for the first time; we have applied CAM theory to obtain a complementary expression for the scattering amplitude (); and we have used this expression to calculate scattering cross sections numerically, showing precise agreement with the numerical results of the partial-wave expansion first obtained in Ref.~\cite{Dolan:2017rtj}.

The spectrum of Regge poles for the neutron star case ($R/M = 6$) exhibits two distinct branches of poles (see ), which we have labelled `broad resonances' and `surface waves' \cite{}. Ultra-compact objects have, in addition, a third branch of `narrow resonances' corresponding to modes inside a cavity between the light-ring and their surface.

Links between QNMs and Regge poles. Further examination.

By applying a WKB method at lowest order, we have obtained a better understanding of the spectrum of the broad resonances, finding that the magnitude of the imaginary part is linked to the magnitude of the discontinuity in the effective potential at the surface of the body (see Eq.~). Taking the WKB method to higher orders will presumably improve the accuracy of approximation. Further work is need to model the surface-wave branch, and the narrow-resonance branch.

We can be confident that there are no additional branches Regge poles in the first quadrant, for two reasons. First, we have searched across the complex-$\lambda$ domain using the scanning method detailed in Sec.~. Second, in our calculation of $d\sigma/d\Omega$, we find a precise numerical agreement between our CAM calculation and our partial-wave calculation; this would not be the case if poles in this domain had been missed.

Complementary method. Not efficient method at present. Convergence of the sum is not as rapid as in the BH case, due to the branch of broad resonances close to the real axis. Neither have we obtained an approximation has powerful as that of Ould Elhadj and Folacci for describing the orbiting phenomenon. In the case of Mie scattering, a powerful resummation approach is that of Regge-Debye poles, reducing to one branch of poles of higher order. This is a possible direction for future work.

We have examined the case of $R=2.26M$, which is above the Buchdahl bound but nevertheless below ... blah blah, work in Tom's text.

Remarkably, in the 1970s, experimental data on rainbows in nucleon-nucleon scattering experiments was used to constrain models of the nuclear potential.

Finally a paragraph on the connections to Mie scattering by a transparent sphere.

\acknowledgments
M. O. E. H. wishes to thank Antoine Folacci for various discussions concerning this work and thanks Sam R. Dolan for his kind
invitation to the University of Sheffield. T.S.~acknowledges financial support from EPSRC. S.R.D.~acknowledges financial support from the European Union's Horizon 2020 research and innovation programme under the H2020-MSCA-RISE-2017 Grant No.~FunFiCO-777740, and from the Science and Technology Facilities Council (STFC) under Grant No.~ST/P000800/1.

>>>>

The Regge pole spectrum for a compact body is qualitatively similar to the Regge pole spectrum found in Mie scattering of electromagnetic waves by a transparent sphere of refractive index $\tilde{n}$. This has been studied since the 1960s; see for example Fig.~9.2 in Ref.~\cite{Nussenzveig:2006}. % In the latter case, the poles are given by the zeros of a combination of special functions.
Henceforth we shall adopt the terminology of Nussenzveig \cite{Nussenzveig:2006}, in which the three branches are labelled as:
\begin{enumerate}
 \item \emph{Broad resonances}: approximately uniformly-spaced poles above the real axis with approximately constant imaginary part; somewhat sensitive to internal structure ($r < R$);
 \item \emph{Surface waves}: highly-damped modes that are relatively insensitive to the internal structure and which depend chiefly on the surface geometry.
 \item \emph{Narrow resonances}: modes approaching the real axis corresponding to trapped modes which only appear in the UCO case ($R < 3M$).
\end{enumerate}
In Figs.~\ref{fig:rp1} and \ref{fig:rp2} the broad resonances, surface waves and narrow resonances are indicated by red squares, blue circles and purple triangles, respectively; and black diamonds indicate the black hole RPs. Whereas there exists an infinite number of poles of the surface-wave branch, in principle, there are no narrow-resonance poles at all in the $R/M = 6$ case, and just $\le 4$ narrow-resonance poles in the $R/M = 2.26$ case. The narrow-resonance branch is observed to end close to the start of the black-hole branch.

@@

Regge poles and quasinormal modes for relativistic stellar models (of which \textit{w}-modes are a sub-category for gravitational perturbations) both satisfy the same wave equation and the same boundary conditions but with different interpretations for the angular momentum index and the frequency. Both types of pole satisfy the regularity condition at the origin~\eqref{bc_1_in} and the condition of a purely outgoing wave in the far field
%
\begin{equation}\label{bc_pole_inf}
\phiout (r) \scriptstyle{\underset{r_\ast \to +\infty}{\sim}}
\displaystyle{  A^{(+)}_{\lambda-1/2} (\omega) e^{+i\omega r_\ast}}.
\end{equation}
%
Thus, the Regge poles are solutions of Eq.~\eqref{H_Radial_equation} for which the  Wronskian of $\phireg$ and $\phiout$ vanishes (\textit{i.e.}, $A^{(-)}_{\lambda_n(\omega)-1/2} (\omega)=0$), viz.,
\begin{equation}\label{Wronskian}
  W[\phireg, \phiout ]= 0 .
\end{equation}
It has been shown \cite{Zhang:2011pq} that the asymptotic expressions of $\phireg$ and $\phiout$ can be derived in asymptotic regions, by using the standard WKB approximation. In the high-frequency limit, the interior solution of Eq.~(\ref{H_Radial_equation}) is (see Ref.~\cite{Zhang:2011pq})
\begin{equation}\label{Approx_origin}
  \phireg \underset{\omega \to \infty}{=} \omega r_* j_{\lambda-1/2} (\omega r_*),  \quad\quad 0\leq r_*\leq R_* ,
\end{equation}
where $j_{\lambda-1/2}(\cdot)$ is the spherical Bessel function of the first kind. In addition
\begin{equation}
\label{Approx_infinity}
\phiout  \scriptstyle{\underset{\omega \to \infty}{=}}
\left\{
\begin{aligned}
&e^{i\omega(r_*-R_*)}+\mathcal{R} e^{-i\omega (r_*-R_*)}& \scriptstyle{1/\omega\leq r_{*} < R_*,}\\
&\left(1+\mathcal{R}\right) e^{i\omega(r_*-R_*)}        & \scriptstyle{R_*\leq r_{*}<\infty.}
\end{aligned}
\right.
\end{equation}
Here $R_*$ is the tortoise coordinate at the surface of the body,
\begin{equation}\label{R_star}
  R_*=\int_{0}^{R}\, dr\,\frac{1}{\sqrt{f(r)h(r)}}
\end{equation}
and $\mathcal{R}$ is a reflection coefficient with the definition given in Ref.~\cite{Berry_1982}.
Because the potential has a direct discontinuity at the surface of the compact body (see. Refs \cite{Zhang:2011pq,Berry_1982} for more details), we have for our model
%
\begin{equation}\label{R_reflection_1}
  \mathcal{R} = \alpha \omega^{-2}
\end{equation}
with
\begin{eqnarray} \label{eq:alpha}
% \nonumber % Remove numbering (before each equation)
  \alpha &=& \frac{1}{4} \Delta V(R) = \pm \frac{3M(R-2M)}{4 R^4} ,
\end{eqnarray}
where $\Delta V$ is the discontinuity in the effective potential at the surface, defined in Eq.~(\ref{def:DeltaV}), and the $+$ ($-$) sign denotes the scalar (axial gravitational wave) case.

@@

In this work, we have: (1) calculated the full spectrum of Regge poles for a compact body spacetime for the first time; (2) applied CAM theory to obtain an expression for the scattering amplitude as a residue series accompanied by a background integral, Eqs.~(\ref{CAM_Scalar_Scattering_amp_tot})--(\ref{CAM_Scalar_Scattering_amp_decomp_RP}); and (3) utilised the CAM formulae to calculated scattering cross sections numerically, demonstrating precise agreement with the numerical results of the partial-wave expansion first computed in Ref.~\cite{Dolan:2017rtj}.

The spectrum of Regge poles exhibits two distinct branches of poles in the neutron star case $R/M = 6$ (see Fig.~\ref{RP_R6_approx_2Mw_3_6_s_1}), which we have labelled `broad resonances' and `surface waves'. Ultra-compact objects have, in addition, a third branch of `narrow resonances' (see Fig.~\ref{RP_R226_approx_2Mw_3_6_s_1}), linked to the existence of an effective cavity between the light-ring and the object's surface. Similar branches of poles arise in scattering by a transparent sphere, and so it is natural to adopt the terminology used to describe these branches in Mie scattering \cite{Nussenzveig:2006}.

By applying a WKB method at lowest order, we have obtained a better understanding of the spectrum of the broad resonances, finding that the magnitude of the imaginary part is linked to the magnitude of the discontinuity in the effective potential at the surface of the body (see Eq.~(\ref{eq:alpha}), Eq.~(\ref{lambda_Approx}) and Fig.~\ref{fig:Veff}). Taking the WKB method to higher orders would improve the accuracy of approximation. Further work is need to obtain asymptotic approximations for the surface-wave branch, and the narrow-resonance branch.

We can be confident that there are no additional branches Regge poles in the first quadrant, for two reasons. First, we have scanned across the complex-$\lambda$ domains shown in Figs.~\ref{RP_R6_approx_2Mw_3_6_s_1} and \ref{RP_R226_approx_2Mw_3_6_s_1}, as detailed in Sec.~\ref{subsec:method}. Second, in our calculation of $d\sigma/d\Omega$, we find a precise numerical agreement between the CAM calculation and the partial-wave calculation; this would not be the case if poles in this domain had been missed.

The CAM toolkit was applied here to gain a complementary practical method for calculating scattering amplitudes. However, in its current form, the method is not particularly efficient, for two reasons; first, it requires the computation of a background integral in addition to residue sum; and second, the poles on the `broad resonance' branch have an approximately uniform spacing $\Delta \lambda \sim 2$, and thus a large number of poles must be found. In the black-hole case, by contrast, there is only a single branch of surface-wave-type modes (see Fig.~\ref{RP_R6_approx_2Mw_3_6_s_1}), and in the high-frequency limit, a single pole captures both the orbiting and glory phenomena \cite{Folacci:2019cmc, Folacci:2019vtt}. In the case of Mie scattering, a powerful extension of the CAM approach is to define the so-called Regge-Debye poles \cite{Nussenzveig:2006}, thereby eliminating the surface-wave branch entirely. This is a possible direction for future investigation.

%In this work, we have  $R=2.26M$, which is above the Buchdahl bound but nevertheless below ... blah blah, work in Tom's text.

%Remarkably, in the 1970s, experimental data on rainbows in nucleon-nucleon scattering experiments was used to constrain models of the nuclear potential.

%Finally a paragraph on the connections to Mie scattering by a transparent sphere.

\acknowledgments
M.O.E.H. wishes to thank Antoine Folacci for various discussions concerning this work and thanks Sam R. Dolan for his kind
invitation to the University of Sheffield. T.S.~acknowledges financial support from EPSRC. S.R.D.~acknowledges financial support from the European Union's Horizon 2020 research and innovation programme under the H2020-MSCA-RISE-2017 Grant No.~FunFiCO-777740, and from the Science and Technology Facilities Council (STFC) under Grant No.~ST/P000800/1.
